### 2012-10-03 1.tex
Edits to 2 and 4, and writing in 3

\chapter{Sequence structure constraints in generative grammar}
\label{msc}

The first part of this text concerns the empirical substance of the theory of \textsc{Morpheme Structure Constraints} (MSCs), that is, language-specific constraints on the shapes of morphs stored in lexical memory.
Evidence is presented to support the claim that such constraints are solely derived by other components of grammar, and do not have an ontology of their own.

The first section of this chapter outlines a brief history of MSCs.
The following section draws upon this material to present the above claim as a null hypothesis, and presents conditions for falsifying it.
The final section outlines the remainder of Part I (chapters 2--4), which present evidence in support of this hypothesis.

% it is not clear what follows from a phonotactic description
% pater on NC
% can't be right tho, see stanley, kisseberth, clayton, shibatani, hooper, etc.

\section{A brief history of MSCs}

\subsection{Structuralism}

Since, with the possible exception of Roman Jakobson (who might be labeled an early generativist), the mentalistic interpretation of structural linguistic analyses was anathema, one must take care to not assume that a description of MSCs is something that these linguists would posit as part of speakers' knowledge of language.

\subsubsection{Morpheme boundaries in phonology}

bloomfield

%While \citeauthor{Bloomfield1930} is known for introducing morphemic structure into phonology, a member of the Prague circle developed an arguably more sophisticated approach only two years later.

\citet{Jakobson1932} attributes a number of phonological properites of the Russian imperative to morpheme juncture. The vast majority of Russian infinitives consist of the stem followed by a theme vowel and /t\pal/. The reflexive is marked wih a final /-sə/ which feeds a general process of regressive voicing assimilation. In the reflexive infinitive, /\ldots t\pal-s\ldots/ is realized [ts], with no palatalization. However, palatalization of a stem-final labial or coronal---including /t\pal/, as in (\ref{russian}c)---is preserved in the reflexive imperative.

\begin{example}
\label{russian}
Russian reflexives \citep[after][]{Jakobson1932}:\footnote{I have taken a number of liberties with \citeauthor{Jakobson1932}'s presentation of the data, which uses an abstract phonemic transcription. Thanks to Lev Blumenfeld (p.c.) for help with the transcription of this data.}

\begin{tabular}{l l l l} %\toprule
   &  infinitive      & imperative \\ %\midrule
a. & [slav\pal itsə]  & [slaf\pal s\pal ə]  & `be glorious'    \\
   & [upram\pal itsə] & [upram\pal s\pal ə] & `be stubborn'    \\
b. & [kras\pal itsə]  & [kras\pal s\pal ə]  & `put on makeup'  \\
   & [ʒar\pal itsə]   & [ʒar\pal s\pal ə]   & `roast'          \\
c. & [zəbytsə]        & [zəbut\pal s\pal ə] & `forget' \\ %\bottomrule
%   & ab\'utsa      & ab\'ujsa     & `put on shoes'   \\
\end{tabular}
\end{example}

\noindent The forms in (\ref{russian}c) have different outcomes for their /t\pal-s/ clusters. \citeauthor{Jakobson1932} proposes that the preservation of palatalization is a special property of the imperative. %A few years later, \citet{Trnka1936}, another member of the Prague Circle, makes the connection between \citeauthor{Jakobson1932}'s hypothesis and phonotactic generalizations.

\subsubsection{The morph as a constraint domain}

A corrolary of this hypothesis is that the phoneme sequences found within morphs may be distinct from those which span multiple morphs.

This was apparent to structuralists

the phonologists 
This is apparent from 
a discussion of constraints on vowel sequences in Mixteco given by \citet{Pike1947b}.
It is apparent that \citeauthor{Pike1947b}'s constraints on vowel sequences are not intended to hold across morph boundaries,

according to the phonological analysis of a Mixteco text published a few years earlier \citep{Pike1944}.

\begin{example}
Mixteco MSCs \citep{Pike1947b} and complex words \citep{Pike1944}:

\begin{tabular}{r l l l} %\toprule
   & MSC & complex exception \\ %\midrule
%a. & *{C}a{C}e & [k\'a-\textsuperscript{n}dee] & `kept \ldots inside' \\
%b. & *{C}\textipa{@}{C}e & [n\`i-k\`ə bə-de] & `who entered'        \\
%c. & *{C}e{C}i & [te-n\'i-ke\textsuperscript{n}da] & `was walking         \\
%d. & *{C}i{C}e & [te-n\`i-kee-t\`ə] & `and went away'      \\ %\toprule
%e. & *{C}e{C}o & b\'e\textglotstopvari e-\v{z}\'o & `our house'          \\
%f. & *{C}eo & ke-o-d\'e & `we eat him'         \\
a. & *{C}a{C}e & [ká\textsuperscript{n}dee] & `kept \ldots inside' \\
b. & *{C}ə{C}e & [nìk\`əbəde] & `who entered'        \\
c. & *{C}e{C}i & [teníke\textsuperscript{n}da] & `was walking         \\
d. & *{C}i{C}e & [tenìkeet\`ə] & `and went away'      \\ 
\end{tabular}
\end{example}

\noindent \citet[][166]{Pike1947b} affirms that the morpheme is ``marked'' by the violation of morpheme-internal sequence restrictions, an idea further developed by \citet{Harris1955} as a morpheme discovery routine.

\subsubsection{Biuniqueness and neutralization}

joos
harris
chomsky

\subsection{Early generativism}

\subsubsection{\citealt{Stanley1967}}

inventory constraints
sequence constraints

\subsubsection{Surface constraints and occultation}

shibatani
stampe
dell

\emph{A}[mt] `office', \emph{Anbau} `cultivation'
\emph{ei}[n.g]\emph{reifen} `to intervene',

\subsubsection{The duplication problem}

\citeauthor{Anderson1974} sees the duplication of \textsc{Backness Harmony} as both sequence structure constraint and phonological rule as a foregone conclusion
.The tendency of sequence structure constraints to duplicate phonological rules was noted by \citet[][401f.]{Stanley1967} and in \emph{SPE} (p.~382) in the discussion of morpheme structure constraints, and is labeled ``the duplication problem'' by \citet[][?]{Kenstowicz1977}.

inventory problems 
dell, chung, but also problems

conspiracies

\subsection{Autosegmentalism}

also ``supersegmentalism''

The proposals during this period are very diverse.

\subsubsection{Syllable structure}

hooper
noske

\subsubsection{Morph features}

Leben
Kaye

\subsubsection{The Obligatory Contour Principle}

\subsection{Classic Optimality Theory}

%(page numbers here are taken from the published version)

\subsubsection{Richness of the Base}

some quotes from OT doc
\citet{PE}
\citet{Bye2001}

\subsubsection{Freedom of Analysis}

\citet{Smolensky1996}
\citet{PE}

\subsubsection{Undominated constraints}

% complaints about maceahern, gallagher
% bit on loanword adaptation

\subsection{Recent developments}

\subsubsection{Language acquisition}
% jusczyk forever
% white et al.
% early word learning
% syllables

\subsubsection{Psycholinguistic tasks in adults}

% wordlikeness
% spotting
% error correction, etc.

\subsubsection{Non-native word adaptation}

% perceptual factors

\section{The null hypothesis}

\subsection{The empirical proposal}

\subsection{The acquisition principle}

\subsection{Falsifiability}

%\subsection{Universal constraints?}    \input{1.2.4}

\section{Outline of the dissertation}

%Chapter 2
%Chapter 3
%Chapter 4

>>>>

\chapter{A generative theory of phonotactics and phonotactic learning}

\section{\emph{Táxis} and generativity}

%\emph{τάξις}
(\emph{táxis} `order, arrangement; the prefix \emph{syn-} means `together')

This etymology gives little clue to just what syntacticians do, however. This is not merely a consequence of the ``generative revolution'', as the predecessors of transformational grammars associated with Zellig Harris and Noam Chomsky were also more than theories of arrangement, but also instantiate the notion of derivation \citep[e.g.,][]{Post1943}.

\subsection{?}

\subsection{Units of description}
\subsubsection{Segment inventory}
\subsubsection{Surface segment sequences}
\subsubsection{Underlying segment sequences}

\subsection{The morph}

The notion of blocking in non-derived environments (\citealp[163]{Kiparsky1973a}, \citeyear[152]{Kiparsky1982a}, \citealp{Mascaro1976}) is well-known, but \citet{Hall2006} also amasses evidence for phonological processes restricted to non-derived environments. One of the most famous examples is the distribution of \emph{ich-laut} [ç] and \emph{ach-laut} [x] in German \citep{Bloomfield1930}. As shown in (\label{ichach}), [x] appears only before back vowels; in all other contexts, the only dorsal fricative is [ç].

\begin{example}[German \emph{ich}- and \emph{ach-laut}]
\begin{tabular}{l l l}
a. & \emph{B}[uːx]          & `book'           \\
   & \emph{T}[ɔx]\emph{ter} & `daughter'       \\
   & \emph{N}[ax]\emph{t}   & `night'          \\
b. & \emph{S}[iç]\emph{t}   & `view'           \\
   & \emph{Sp}[eç]\emph{t}  & `woodpecker'     \\
   & \emph{Ger}[yç]\emph{t} & `rumor'          \\
   & \emph{Kn}[øç]\emph{el} & `ankle, knuckle' \\
   & \emph{Fl}[ɛç]\emph{e}  & `surface'        \\
\end{tabular}
\end{example}

%a. & \emph{B}[uːx]          & \emph{B}[yːç]\emph{er} & `book' \\
%   & \emph{L}[ɔx]           & \emph{L}[øç]\emph{er}  & `hole' \\
%%   &\emph{T}[ɔx]\emph{ter} & \emph{T}[øç]\emph{ter}
%   & \emph{N}[ax]\emph{t}   & \emph{N}[ɛç]\emph{te} & `night' \\

\emph{K}[uːx]\emph{en}
\emph{K}[uːç]\emph{en}

%\emph{Mas}[oːx]-\emph{Mas}
\footnote{Examples like \emph{Mas}[oːç]\emph{ist} `masochist', \emph{Eun}[uːç]\emph{ismus} `eunichism', first noted by \citet{Merchant1994}, suggest that this should also be restricted to assimilation within the same foot \citep[226f.]{Jensen2000}.}

%\emph{Mas}[oːx]-\emph{Mas}[oːç]\emph{ist}

The concept of non-derived environment blocking

\section{The model}

\subsection{Prosodic parsing}
\subsubsection{Syllabification}
\subsubsection{Autosegmental tiers}

\subsection{Rule application}
\subsubsection{Violation detection}
\subsubsection{Rule exceptionality}

\section{Consequences of the model}

\subsection{Gradience}
\subsubsection{Probabilistic well-formedness}
\subsubsection{Cumulativity}
\subsubsection{External factors}

\subsection{Static phonotactics}
\subsubsection{Statistical insensitivity}
\subsubsection{Saussurean arbitrariness}

\subsection{Order of acquisition}
\subsubsection{Lexical acquisition}
\subsubsection{Prosodic parsing}
\subsubsection{Rule application}

%\subsubsection{Exceptionality}
